---
title: "Case study: scoring the CDISC pilot ADaM package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: scoring the CDISC pilot ADaM package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_pkgs <- all(vapply(
  c("metacore", "pharmaverseadam", "r4subscore"),
  requireNamespace, logical(1), quietly = TRUE
))
knitr::opts_chunk$set(eval = has_pkgs)
```

This case study scores a real, public ADaM package end to end. Nothing is
hand-tuned: we take two artifacts that were produced **independently** and ask
how ready the result is for submission.

- The **specification** comes from
  [metacore](https://github.com/pharmaverse/metacore): the CDISC pilot ADaM
  metadata, describing what each dataset and variable should be.
- The **datasets** come from
  [pharmaverseadam](https://github.com/pharmaverse/pharmaverseadam): ADaM data
  built with [admiral](https://github.com/pharmaverse/admiral).

Because the spec and the data were not made to match each other, the score is
not a formality. It tells us where the package stands.

```{r, eval = TRUE, echo = FALSE, results = "asis"}
if (!has_pkgs) {
  cat("> The packages needed to run this case study (metacore,",
      "pharmaverseadam, r4subscore) are not installed, so the code below is",
      "shown but not evaluated.\n")
}
```

## The inputs

```{r setup}
library(r4subpharma)

# The CDISC pilot ADaM specification (covers ADSL, ADAE, ADLBC, ADADAS, ADTTE)
e <- new.env()
load(metacore::metacore_example("pilot_ADaM.rda"), envir = e)
spec <- e$metacore

# One of the admiral-built datasets
adsl <- pharmaverseadam::adsl
dim(adsl)
```

## One call to a score

`submission_readiness()` harvests documentation evidence from the spec and
conformance evidence from the dataset, then computes the Submission Confidence
Index.

```{r score-adsl}
ctx <- r4subcore::r4sub_run_context("CDISCPILOT01", "PROD")
spec_adsl <- metacore::select_dataset(spec, "ADSL", verbose = "silent")

res <- submission_readiness(list(ADSL = adsl), spec_adsl, ctx)
res
```

## Reading the score

The pillar breakdown shows where the number comes from.

```{r pillars}
as.data.frame(res$sci$pillar_scores)
```

Quality (documentation and types) and usability (labels) are strong. The weak
pillar is **traceability**: the specification lists variables that the built
dataset does not yet contain. That single gap is what holds the score back.

## The gap, made actionable

The value of a score is the fix list behind it. The failing traceability checks
name exactly which specified variables are missing from the data.

```{r gaps}
ev <- res$evidence
missing <- ev[ev$indicator_id == "T-ADAM-001" & ev$result == "fail", ]
nrow(missing)
sub(".*: ", "", missing$message)
```

These are the CDISC pilot analysis variables: treatment-coded flags, age and
site groupings, and completion flags. An analysis-ready ADSL is expected to
carry them, so their absence is a real readiness finding, not a formatting nit.

We can also see the smaller quality signals: any variable whose type does not
match the specification.

```{r types}
ev[ev$indicator_id == "Q-ADAM-001" & ev$result == "warn",
   c("location", "message")]
```

## Scoring more of the package

The same call scales to several datasets at once; the score then reflects the
whole set of evidence.

```{r package}
res_pkg <- submission_readiness(
  list(ADSL = pharmaverseadam::adsl, ADAE = pharmaverseadam::adae),
  spec,
  ctx
)
res_pkg$sci$SCI
as.data.frame(res_pkg$sci$pillar_scores)
```

## Takeaways

- The score is honest. On genuinely independent inputs it lands in the
  "minor gaps" band, driven down by a specific, explainable traceability gap.
- The evidence is actionable: the failing checks are a prioritized list of
  variables to add, not an opaque number.
- It is reproducible on public data, and the evidence table flows into the rest
  of the R4SUB ecosystem. Use `r4subprofile` for authority-specific weighting and
  `r4subrisk` for risk quantification, with no extra work.
```
